Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2024-08-01 doc/manuscript/response-to-reviews.Rmd

*(NEEDS RESPONSE. We did not receive this information directly. We will have to figure out which records are missing)*

Another example, the map (Fig 2) is fine as an illustration or proof of the global participation to the efforts of reporting, but it doesn’t help identify the gaps in the current reporting. Some vegetation types are well represented, others have a serious deficit. Identifying and quantifying this, for instance in terms of report per area, would have been a more eloquent and useful than a map. The imbalance in the data goes for the below/aboveground compartments too.

**We have made some of the suggested revisions (detailed below) to improve the assessment of data representativeness and consider this to be an important improvement. At the same time, we also wish to emphasize that the goal of this paper was to document the process and provide an overview of the types of data submitted, as opposed to submitting data that adequately represents all forest types and documenting their representativeness. The latter -- to the extent that it would even be possible -- would be a massive project far beyond the scope of this paper. Therefore, while improving the assessment of data representativeness, we believe it would be overkill to analyze all aspects of representativeness.**

**We have added a graph to the map (Fig. `r fig_map`) showing the proportional breakdown by FAO ecozone of (1) area with tree cover, (2) sites and records submitted to EFDB, (3) sites and records submitted to EFDB. We have also added a statement on relative representation by biome: "Relative to forested area, submitted data under-represented the tropics and over-represented temperate regions, mirroring the composition of ForC (Fig. `r fig_map`)."**

**To assess the data distribution within in the global climate space, we have added a Whittaker-style diagram showing ForC sites and those with submitted data within the MAT-MAP climate space (new Fig. `r fig_climate`). We have also added a statement on climate representativity: "Submitted data covered all forested climates with substantial land area except the colder boreal regions (boreal tundra woodland; Figs. `r fig_map`, `r fig_climate`) and were concentrated in climates that are relatively common over Earth's land area (Fig. `r fig_climate`)." **

@@

**Regarding imbalance among variable types, `r table_variables` provides a complete record of the numbers of records in ForC and submitted to EFDB. We feel that further analyzing by categories such as ecozone or age would be overkill, make the manuscript too long, and of little interest to most readers.**

Fig. 3 *(now 4)* isn’t conclusive either because ecozones have different sizes, so the number of reports on itself is not a sufficient to evaluate the coverage. The representativity of the data within an ecozone depends on the variability between records: if they all cover the same age classes, again, the representativity is bad.

**To evaluate the coverage, we have modified this figure to include bars representing the relative area (% of global) of each category and added some text summarizing the results/ citing this figure (3rd paragraph in Results).**

**Again, we feel that deeper analysis (e.g., age x ecozone) of the representativity would be overkill and of little interest to most readers.**

### 2024-08-01 doc/manuscript/ForC-EFDB manuscript.Rmd
add figures and text responding to reviewer 1 comments
#63 , #62 , #61

---

@@

ForC v4.0 records and the subset submitted to EFDB were distributed across all forested continents, biomes, and forest types, albeit very unevenly (Figs. `r fig_map`-`r fig_histograms`).
Relative to forested area, submitted data under-represented the tropics and over-represented temperate and boreal biomes (Fig. `r fig_map`).
Submitted data were concentrated in climates that are relatively common over Earth's land area (Fig. `r fig_climate`).
Among the records submitted to EFDB, the largest number came from North America, followed by Asia, South America, and Africa (Fig. `r fig_histograms`c), with the most represented FAO ecozones being boreal coniferous forest, temperate continental forest, and temperate mountain systems, followed by tropical rain forests and moist deciduous forests (Fig. `r fig_histograms`b).
By far the most records came from needleleaf evergreen forests, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`b). 
The largest number of records came from mature forests (>100 years), followed by young and intermediate-aged stands (Fig. `r fig_histograms`d).

@@

![**Figure `r fig_map`. Map of sites in ForC shaded by number of independent records relevant (circles) and submitted (triangles) to EFDB.**  Underlying map shows FAO ecozones, and symbols are colored according to the number of records at each site. Inset stacked bar chart shows proportional representation by FAO ecozone for tree cover area (from SYNMAP,@jung_exploiting_2006, relevant ForC records, and submitted records. FAO ecozones are coded as follows: Ba-Boreal coniferous forest, Bb-Boreal tundra woodland, BM-Boreal mountain systems, P-Polar, SBSh-Subtropical steppe, SBWh-Subtropical desert, SCf-Subtropical humid forest, SCs-Subtropical dry forest, SM-Subtropical mountain systems, TAr-Tropical rain forest,  TAwa-Tropical moist deciduous forest, TAwb-Tropical dry forest, TBSh-Tropical shrubland, TBWh-Tropical desert, TeBSk-Temperate steppe, TeBWk-Temperate desert, TeDc-Temperate continental forest, TeDo-Temperate oceanic forest, TeM-Temperate mountain systems, TM-Tropical mountain systems.](figures_tables/World_Map_of_sites_with_FAO_and_IPCC_data_sent.png)

![**Figure `r fig_climate`. Distribution of ForC records relevant (dots) and submitted (triangles) to EFDB within the global climate space of mean annual temperature (MAT) and mean annual precipitation (MAP).** Climate data are from *WorldClim 2 (0.5 resolution, Fick & Hijmans, 2017*. Background colors indicate the global land area with each MAT-MAP combination. GitHub issue 61:https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.15353](figures_tables/Climate_of_forC_sites_IPCC.png)

![**Figure `r fig_histograms`. Histograms of number of independent records in ForC relevant to (grey) and submitted to (black) EFDB, organized by (a) dominant vegetation type, (b) FAO ecozone, (c) continent, and (d) stand age.** For dominant vegetation (a), 'Other' includes deciduous needleleaf, mixed broadleaf- needleleaf, non-woody vegetation (e.g., early successional), and incompletely classified or mixed forest types. For FAO ecozones (b), codes are as listed in the caption of Figure `r fig_map`.](figures_tables/Histogram_n_Relevant_and_Transferred_Records.png)

>>>>

nocite: |
  @harris_updated_2014; @jung_exploiting_2006; @besnard_mapping_2021
---

@@

ForC v4.0 records and the subset submitted to EFDB were distributed across all forested continents, biomes, and forest types, albeit unevenly (Figs. `r fig_map`-`r fig_histograms`).
Relative to forested area, submitted data under-represented the tropics and over-represented temperate regions, mirroring the composition of ForC (Figs. `r fig_map`, `r fig_histograms`b).
Submitted data covered all forested climates with substantial land area except the colder boreal regions (boreal tundra woodland; Figs. `r fig_map`, `r fig_climate`) and were concentrated in climates that are relatively common over Earth's land area (Fig. `r fig_climate`).
Among the records submitted to EFDB, the largest number came from North America, followed by Asia, South America, and Africa, with strongly disproportionate representation of North America and under-representation of South America, Africa, Europe, Australia, and Oceania (Fig. `r fig_histograms`c).
In terms of FAO ecozones, boreal coniferous forest, temperate continental forest, and temperate mountain systems had the most records and were the best represented relative to their global area, whereas other ecozones were poorly represented relative to their global area (Fig. `r fig_histograms`b).
By far the most records and best representation relative to global area were for needleleaf evergreen forests, followed by broadleaf deciduous and broadleaf evergreen (Fig. `r fig_histograms`b). 
The largest number of records came from mature forests (>100 years), followed by young and intermediate-aged stands, but young (<20 year) stands were best represented relative to their global area (Fig. `r fig_histograms`d).

@@

![**Figure `r fig_map`. Map of sites in ForC shaded by number of independent records relevant (circles) and submitted (triangles) to EFDB.**  Underlying map shows FAO ecozones, and symbols are colored according to the number of records at each site. Inset stacked bar chart shows proportional representation by FAO ecozone for tree cover area (from SYNMAP, Jung *et al*, 2006), relevant ForC records, and submitted records. FAO ecozones (obtained from FAO’s GEOnetwork: http://www.fao.org:80/geonetwork) are coded as follows: Ba-Boreal coniferous forest, Bb-Boreal tundra woodland, BM-Boreal mountain systems, P-Polar, SBSh-Subtropical steppe, SBWh-Subtropical desert, SCf-Subtropical humid forest, SCs-Subtropical dry forest, SM-Subtropical mountain systems, TAr-Tropical rain forest,  TAwa-Tropical moist deciduous forest, TAwb-Tropical dry forest, TBSh-Tropical shrubland, TBWh-Tropical desert, TeBSk-Temperate steppe, TeBWk-Temperate desert, TeDc-Temperate continental forest, TeDo-Temperate oceanic forest, TeM-Temperate mountain systems, TM-Tropical mountain systems.](figures_tables/World_Map_of_sites_with_FAO_and_IPCC_data_sent.png)

![**Figure `r fig_climate`. Distribution of ForC records relevant (dots) and submitted (triangles) to EFDB within the global climate space of mean annual temperature (MAT) and mean annual precipitation (MAP).** Climate data are from the CRU TS3.10 dataset (0.5° resolution, 1990-2014, Harris *et al*, 2014). Background colors indicate the global land area with each MAT-MAP combination.](figures_tables/Climate_of_forC_sites_IPCC.png)

![**Figure `r fig_histograms`. Histograms of number of independent records in ForC relevant (grey) and submitted (black) to EFDB (both left axis), along with the relative global area of each type (red, right axis), organized by (a) dominant vegetation type, (b) FAO ecozone, (c) continent, and (d) stand age.** For dominant vegetation (a), 'Other' includes deciduous needleleaf, mixed broadleaf- needleleaf, non-woody vegetation (e.g., early successional), and incompletely classified or mixed forest types. Global coverage of each was obtained from SYNMAP (Jung *et al*, 2006). For FAO ecozones (b), codes are as listed in the caption of Figure `r fig_map`. The relative area of forests by age class was obtained from the global forest age database of Besnard *et al.* (2021), updated in 2024 (DOI: xxx).  ](figures_tables/Histogram_n_Relevant_and_Transferred_Records.png)
